---
title: "Adjusting max size of the Final Course Model"
format: pdf
editor: visual
---

## Load in Libraries

```{r}
#| results: 'hide'
#| message: false
#| warning: false
library(mizer)
library(mizerExperimental)
library(tidyverse)
library(rfishbase)
```

## Load in Course Model

```{r}
#| label: loading final course model
download.file("https://github.com/gustavdelius/mizerCourse/raw/master/build/cel_model_landings.rds",
              destfile = "cel_model_landings.rds")
cel_model_course <- readParams("cel_model_landings.rds")
gear_params(cel_model_course)$yield_observed <- readRDS("celtic_yields.rds")
```

Making a data frame with the new w_max and l_max values

```{r}
cel_model_course <- readParams("cel_model_landings.rds")
params<-species_params(cel_model_course)
saveRDS(params, file = "species_params.rds")
params <- readRDS("species_params.rds")

download.file("https://github.com/gustavdelius/mizerCourse/raw/master/build/species_params_spence_et_al.rds",
              destfile = "species_params_spence_et_al.rds")
sp_spence <- readRDS("species_params_spence_et_al.rds")
sp <- select(sp_spence, species)
sp$latin_name <- c("Clupea harengus", # Herring
                   "Sprattus sprattus", # Sprat
                   "Gadus morhua", # Cod
                   "Melanogrammus aeglefinus", # Haddock
                   "Merlangius merlangus", # Whiting
                   "Micromesistius poutassou", # Blue whiting
                   "Trisopterus esmarkii", # Norway Pout
                   "Trisopterus minutus", # Poor Cod
                   "Merluccius merluccius", # European Hake
                   "Lophius piscatorius", # Monkfish
                   "Trachurus trachurus", # Horse Mackerel
                   "Scomber scombrus", # Mackerel
                   "Limanda limanda", # Common Dab
                   "Pleuronectes platessa", # Plaice
                   "Lepidorhombus whiffiagonis", # Megrim
                   "Solea solea", # Sole
                   "Capros aper") # Boarfish
max_size_fishbase <- rfishbase::species(sp$latin_name) |>
    select(latin_name = Species, w_max = Weight, l_max = Length)
max_size_fishbase <- left_join(sp, max_size_fishbase, by = "latin_name")
length_weight <- estimate(max_size_fishbase$latin_name, fields = c("Species", "a", "b"))
max_size <- max_size_fishbase |>
    left_join(length_weight, by = c("latin_name" = "Species"))
max_size <- max_size |>
    mutate(w_max_calc = a * l_max ^ b)
max_size$w_max[is.na(max_size$w_max)] <- max_size$w_max_calc[is.na(max_size$w_max)]
max_size<-max_size[c("species","w_max","l_max")]
```

```{r}
#new species parameters
max_size

# Merge new w_max and l_max values
cel_model_course@species_params <- merge(
  cel_model_course@species_params, 
  max_size, 
  by = "species", 
  suffixes = c("", ".new")
)

# Overwrite old values
cel_model_course@species_params$w_max <- cel_model_course@species_params$w_max.new
cel_model_course@species_params$l_max <- cel_model_course@species_params$l_max.new

# Clean up the merged columns
cel_model_course@species_params <- cel_model_course@species_params[, !grepl(".new$", names(cel_model_course@species_params))]

#validate
#cel_model_course <- validParams(cel_model_course)
```

Find an error, due to the changing of w_max, which occured due to the fact the w_max now exceeds the model's maximum grid weight and that now w_max is internally inconsistent with other various slots such as psi, intake_max, etc.

To fix this we first need to do some diagnosis of the problem and check if the new w_max value extends past the weight grid

```{r}
max(cel_model_course@species_params$w_max)
max(cel_model_course@w)
```

Yes it does extend past the weight grid: w_max is \> than w, therefore we need to adjust the weight grid.

```{r}
# Save updated species_params before rebuilding
updated_species_params <- cel_model_course@species_params

# Rebuild MizerParams with updated species_params and same other settings
cel_model_course <- newMultispeciesParams(
  species_params = updated_species_params,
  interaction = cel_model_course@interaction,
  gear_params = cel_model_course@gear_params,
  no_w = 100,  # Or adjust for finer resolution
  min_w = min(cel_model_course@w),
  max_w = max(updated_species_params$w_max) * 1.1  # 10% buffer
)
#validObject(cel_model_course)
saveRDS(cel_model_course, "cel_model_course_updated.rds")
```

Lets check the new model and run to a steady state in tuneParams

```{r}
cel_model_course_updated<-readParams("cel_model_course_updated.rds")
cel_model_course_updated<-tuneParams(cel_model_course_updated)
saveParams(cel_model_course_updated, "cel_model_course_updated.rds")
```

After clicking steady twice the model does not converge after 99 years and requires an unrealistic erepro greater than 1 for Blue whiting, Boarfish, Cod, Common Dab, European Hake, Haddock, Herring etc... additionally the feeding level is at 1 for all species.
